Apache Storm

### 2017-11-30 paper.tex
Added Concepts of Apache Storm

Data can be considered the oil of the current age, because with huge amounts of data one can determine hidden patterns. A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover something new. The question on how to classify any data as big data can be answered by using the 3Vs of big data: Volume, Variety and Velocity. Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data. In this paper regarding \textbf{Apache Storm} we consider the third V of big data to be the influencing factor.

@@

>>>>

Data can be considered the oil of the current age, because with huge amounts of data one can determine hidden patterns. A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover something new. The question on how to classify any data as big data can be answered by using the 3Vs of big data: Volume, Variety and Velocity [2]. Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data. In this paper regarding \textbf{Apache Storm} we consider the third V of big data to be the influencing factor.

@@

Apache storm is one of the available open source tools for stream data processing, which is considered to be one of the most easy to use, fast and fault tolerant systems available [3]. These three factors have been the major points for the general public to choose \textbf{Apache Storm} above other available tools. It was originally created by Nathan Marz at BackType and till now, at present more than 60 companies use or experiment with this tool.

\subsection{Architecture}
The basic Storm data processing architecture consists of stream of tuples flowing through a defined topology to yield the required analytic result. Processing of this stream data is performed in Nodes called the worker nodes, which are responsible for performing the required calculations. There is also another node named the Master Node, which is responsible for the complete Storm infrastructure,  by keeping track of all the worker nodes and the total topology itself.

\begin{figure}
  \begin{center}
    \includegraphics[width=.5\textwidth]{arch.png}
    \caption{High Level Storm Architecture}
    \label{fig:arch}
   \end{center}
\end{figure}

\subsubsection{Nimbus}

\paragraph{The Master node} also called the \textbf{Nimbus} is also responsible for distribution and coordination of the execution of topologies. The Nimbus is the first point where the user submits the defined topology, which is then distributed among the worker nodes.  Figure \ref{fig:arch} illustrates the high level architecture where, Nimbus monitors all the available worker nodes and both the nimbus as well as the worker nodes use \textbf{Apache Zookeeper} to store their states. The actual processing is done by the \textbf{worker nodes}, and each worker node runs multiple Worker Processes each of them mapped to their own topology [4]. It is also worth while noting that multiple worker processes in a worker node could be executing different parts of the topology.

\subsubsection{Worker Node}
\paragraph{Worker Node} is responsible for all the heavy lifting required for processing of the topology. One worker node may run a multiple number of worker processes. Each of these worker processes may belong to the same topology or different topologies. A worker node maintains a \textbf{Supervisor} and a multiple number of Worker Processes which in turn maintains a multiple number of executors. Thus, Executors enable intra-topology parallelism whereas Tasks provide intra-spout or intra-bolt parallelism. Figure \ref{fig:workerarch} depicts the design of a Worker node.[4]

\begin{figure}
  \begin{center}
    \includegraphics[width=.5\textwidth]{worker.png}
    \caption{Worker Node Architecture}
    \label{fig:workerarch}
   \end{center}
\end{figure}

\paragraph{Supervisor} is running in every Worker Node and communicates with the Nimbus to receive assignments from it. Apart from taking assignments from the Nimbus it is also responsible for monitoring the health of the of the worker processes and respawns the worker processes in case it dies.

\paragraph{Worker Process} is the Entity that does the processing and is mapped to a certain topology. Each worker node spawns a JVM and runs one or multiple executors. These executors may be running any of the bolt or spout within the topology, thus providing an added level of parallelism.

\paragraph{Executors} are made up of one or multiple Tasks. Unless specified explicitly, storm assigns single task to a single executor. Executors takes tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue [4].

\subsubsection{Topology}
\paragraph{Topology} is a directed graph where the vertices represent computation and the edges represents the data flow.
